---
title: "Analysis of Santiago Urban Tree by Land Use"
author: "spp"
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

Script to fit Biomass model from Urban trees in Santiago.

####Get Data
```{r Reading Data, eval=FALSE}
#Load Packages
require(dplyr)
require(gdata)

#Set working directory
#setwd("~/Google Drive/Urban")
#For Windows
#setwd("C:/Users/Sebastian/Google Drive/Urban")

# #Maybe a code for switching between linux and windoews
if (Sys.info()['sysname'] == 'Windows'){
  setwd("C:/Users/Sebastian/Google Drive/Urban")
} else if (Sys.info()['sysname'] == 'Linux'){
  setwd("~/Google Drive/Urban")
}

source("Analysis/helpers.R")   

#list of trees in each plot
arboles <- read.xls(xls="Data/arboles_2002-2014.xls", sheet = 'Parcelas con arboles' )
##list of trees present in both dates. Better used by exported csv arboles_growth
#spp.both <- read.xls(by_tailnum <- group_by(flights, tailnum)xls="Data/arboles_2002-2014.xls", sheet = 'Spp in both yrs')

#Species parameters dataframe
species <- read.csv("Data/species.csv")
rownames(species) <- species$Specie


#detailed measurements of each tree in 2014 plots. Height to Crown. Tree height, % shrub below
#TreeBiom <- read.xls(xls="Data/TreeBiometrics.xls", sheet = 'Trees')

#percent of plot used in the different classes. ORMITP?
#FieldLandUses <- read.xls(xls="Data/FieldLandUses.xls", sheet = 'FieldLandUses')

#Coordinates, Land use ID on MAP, Date sampled, crew, address, groundcovers from PlotGrndCvr.xls
Plots <- read.csv("Data/Plots_original.csv")
rownames(Plots) <- Plots$Plot.ID
#Land use ID used on table above. Description and Abbrev. Sumn of area sampled by Land use
LandUseID <- read.csv("Data/PlotLandUSe_Code.csv")
rownames(LandUseID ) <- LandUseID$Land.Use.ID

#Description of shrubs in each plot. 
#Shrubs <- read.xls(xls="Datby_tailnum <- group_by(flights, tailnum)by_tailnum <- group_by(flights, tailnum)a/Shrubs.xls", sheet = 'Shrubs')

#Percent of tree, shrubs and plantables in each plot 
#TreeShrbCvr <- read.xls(xls="Data/TreeShrbCvr.xls", sheet = 'Subplots')
```

####Biomass estimation and definition of allometric equations
1: $\exp(a + b \log(\mathrm{DBH}))$

2: $a + b\mathrm{DBH} + c\mathrm{DBH}^d + e\mathrm{H}^f + g\mathrm{DBH}^2\mathrm{H}$

3: $a \pi(\mathrm{Crown} / 2)^2 \times b \mathrm{H}^c$

4: $a\mathrm{H}^b(\mathrm{DBH}/c)^d$

5: $a\mathrm{DBH}^b+c\mathrm{DBH}^d$

6: $\exp(a+b\mathrm{DBH})$

7: $a+b\log(\mathrm{DBH})$

0: No Formula

####Data preparation
```{r Data preparation, eval=FALSE}

#Biomass estimation from formulas using helper function
arboles$Biomass.2002 = biomass(spp = arboles$Specie,
                               dbh =arboles$DBH2002,
                               h = arboles$H2002,
                               crown = arboles$CROWN2002)
arboles$Biomass.2014 = biomass(spp = arboles$Specie,
                               dbh =arboles$DBH2014, 
                               h = arboles$H2014,
                               crown = arboles$CROWN2014)


#Change to NA the values of DBH2014, H2014 and CROWN2014 from the removed trees based on Removed.2014 == 1
arboles$DBH2014[arboles$Removed.2014==1]  = NA
arboles$H2014[arboles$Removed.2014==1]  = NA
arboles$CROWN2014[arboles$Removed.2014==1]  = NA
arboles$Biomass.2014[arboles$Removed.2014==1]  = 0
arboles$Differences.in.growth = NULL
arboles$Biomass14.02 = NULL  #Eliminates previously biomass estimates
arboles$DBH2002b = NULL #Eliminates weird second dbh
arboles$DBH2014b = NULL #Eliminates weird second dbh

#Change to NA the values of DBH2002, H2002 and CROWN2002 from the ingrowth trees based on In.growth.2014 == 1
arboles$DBH2002[arboles$In.growth.2014==1]  = NA
arboles$H2002[arboles$In.growth.2014==1]  = NA
arboles$CROWN2002[arboles$In.growth.2014==1]  = NA
arboles$Biomass.2002[arboles$In.growth.2014==1]  = 0 #Serves to properly estimat e change later.

#Do we need to estimate growth on 10 years?
#arboles$DBH2012 = arboles$DBH2002  + (arboles$DBH2014 - arboles$DBH2002)/ arboles$Days.between.sampling.dates * (365 * 10)
#arboles$H2012 = arboles$H2002  + (arboles$H2014 - arboles$H2002)/ arboles$Days.between.sampling.dates * (365 * 10)


#The change in DBH column only has the trees with measurments in both times. #Only for growth regression
#Considering a linear dbh growth
arboles$DBH.02.12 = (arboles$DBH2014 - arboles$DBH2002) / arboles$Days.between.sampling.dates * (365 * 10)
arboles$DBH.02.12[arboles$In.growth.2014==1]  = NA 

#Estimate basal area in square meters
arboles$BA2002 = (arboles$DBH2002/2)^2 * pi /10000 
arboles$BA2014 = (arboles$DBH2014/2)^2 * pi /10000
#The change in basal area adds the basal area of the recruited trees.
#considering a linear basal area growht (not estimated with adjusted dbh)
arboles$BA2002[arboles$In.growth.2014==1]  = 0
arboles$BA2014[arboles$Removed.2014==1]  = 0
arboles$BA.02.12 = arboles$BA2014 - arboles$BA2002 / arboles$Days.between.sampling.dates * (365 * 10)
arboles$BA.02.14 = arboles$BA2014 - arboles$BA2002
arboles$BA2002[arboles$In.growth.2014 == 1] = NA
arboles$BA2014[arboles$Removed.2014==1]  = NA

#The change in biomass adds the basal area of the recruited trees.
arboles$Biomass.02.14 = arboles$Biomass.2014 - arboles$Biomass.2002
arboles$Biomass.2002[arboles$In.growth.2014==1] = NA
arboles$Biomass.2014[arboles$Removed.2014==1]  = NA

#Adding Plot land use based on PLot ID to arboles list
arboles$Land.Use.ID = Plots[as.character(arboles$Plot.ID),'Land.Use.ID.on.Map']
arboles$Land.Use.Description = LandUseID[as.character(arboles$Land.Use.ID),'LandUseAbbreviation']
arboles$Land.Use.Description_grouped = LandUseID[as.character(arboles$Land.Use.ID),'LandUseAbbreviation_grouped']
arboles$GrndCvrClass = Plots[as.character(arboles$Land.Use.ID),'Perm_degree']

#Add species code and classification of trees from excel
arboles$Class = species[as.character(arboles$Specie),'Class']
arboles$Specie_code = species[as.character(arboles$Specie),'Specie_code']

#reduce and export the database if growth trees to avoid problems
write.csv(arboles[!is.na(arboles$DBH.02.12),] , file = 'Data/arboles_growth.csv', row.names = FALSE)

#reduce and export the database if growth trees to avoid problems
#write.csv(arboles , file = 'Data/arboles_2002-2014_2.csv', row.names = FALSE)
#Code to plot a histogram of biomass to detect outliers
#hist(arboles$Biomass.2002, xlim = c(0, 500))
```

#### By Plot statistics table export
```{r Plot Statistics, eval=FALSE}
require(dplyr)
by_plot <- dplyr::group_by(arboles, Plot.ID)
by_plot_results = summarise (by_plot,
                             Land.Use.Description = unique(Land.Use.Description),
                             Land.Use.Description_grouped = unique(Land.Use.Description_grouped),
                             GrndCvrClass  = unique(GrndCvrClass), 
                             num_species = length(unique(Specie)),
                             num_trees_2002 = length(Specie) - sum(In.growth.2014, na.rm = TRUE),
                             num_trees_2014 = length(Specie) - sum(Removed.2014, na.rm = TRUE),
                             removed_trees.2014 = sum(Removed.2014, na.rm = TRUE),
                             ingrowth_trees.2014 = sum(In.growth.2014, na.rm = TRUE),
                             BA2002 = sum(BA2002, na.rm = TRUE),
                             BA2014 = sum(BA2014, na.rm = TRUE),
                             BA.02.14 = sum(BA.02.14, na.rm = TRUE),
                             Biomass.2002 = sum(Biomass.2002, na.rm = TRUE),
                             Biomass.2014 = sum(Biomass.2014, na.rm = TRUE),
                             Biomass.02.14 = sum(Biomass.02.14, na.rm = TRUE)
                             )


#plots ya tiene las columnas aquí calculadas. 
#Hay que eliminarlas antes si se quiere exportar de nuevo
#Adding add these results to Plots dataframe and export it for later analysis
Plots = merge(Plots, by_plot_results, by = 'Plot.ID',
              all = TRUE)     #Even if there is no matching, export the line
#Export Plots dataframe to speed later analysis

#Convert NA to zeroes for later analysos
Plots$Land.Use.Description[is.na(Plots$Land.Use.Description)] <- 0
Plots$Land.Use.Description_grouped[is.na(Plots$Land.Use.Description_grouped)] <- 0
Plots$GrndCvrClass[is.na(Plots$GrndCvrClass)] <- 0
Plots$num_species[is.na(Plots$num_species)] <- 0
Plots$num_trees_2002[is.na(Plots$num_trees_2002)] <- 0
Plots$num_trees_2014[is.na(Plots$num_trees_2014)] <- 0
Plots$removed_trees.2014[is.na(Plots$removed_trees.2014)] <- 0
Plots$ingrowth_trees.2014[is.na(Plots$ingrowth_trees.2014)] <- 0
Plots$BA2002[is.na(Plots$BA2002)] <- 0
Plots$BA2014[is.na(Plots$BA2014)] <- 0
Plots$Biomass.2002[is.na(Plots$Biomass.2002)] <- 0
Plots$BA.02.14[is.na(Plots$BA.02.14)] <- 0
Plots$Biomass.02.14[is.na(Plots$Biomass.02.14)] <- 0
Plots$Biomass.2014[is.na(Plots$Biomass.2014)] <- 0

#Export plots analysis to csv file
write.csv(Plots, file = 'Data/Plots.csv', row.names = FALSE)
```


## By Plots statistical analysis and plots
```{r Reading Results Data, eval=TRUE, message=FALSE, warnings = FALSE}
require(ggplot2)
require(gplots)
require(multcomp)

#Set working directory
setwd("~/Google Drive/Urban")
#For Windows
#setwd("C:/Users/Sebastian/Google Drive/Urban")

#Reading Data
Plots <- read.csv("Data/Plots.csv")

#print number of plots per land use (After combining the many land usesws)
table(Plots$Land.Use.Description_grouped)

#======================================================================================
# Anova of BA change, biomass change by land use (R, C/I, G, A, T)

# 2012-2014 Change in Basal Area
fitBA_change <- aov(BA.02.14 ~ Land.Use.Description_grouped , data=Plots)
anova(fitBA_change)

# 2012-2014 Change in Biomass
fitBiomass_change <- aov(Biomass.02.14 ~ Land.Use.Description_grouped , data=Plots)
anova(fitBiomass_change)

#======================================================================================
# Anova of BA, biomass, number of species, mortality
# by grouped land use Classification (R, C/I, G, A, T)

# 2002 Basal Area
fitBA.2002  = aov(BA2002 ~ Land.Use.Description_grouped , data=Plots)
anova(fitBA.2002)          #NOTHING SIGNIFICANT

# 2014 Basal Area
fitBA.2014  = aov(BA2014 ~ Land.Use.Description_grouped , data=Plots)
anova(fitBA.2014)          #NOTHING SIGNIFICANT

# 2002 Biomass
fitBiomass.2002  = aov(Biomass.2002 ~ Land.Use.Description_grouped , data=Plots)
anova(fitBiomass.2002)          #NOTHING SIGNIFICANT

# 2014 Basal Area
fitBiomass.2014  = aov(Biomass.2014 ~ Land.Use.Description_grouped , data=Plots)
anova(fitBiomass.2014)          #NOTHING SIGNIFICANT

# Total number of species
fitSpecies  <- aov(num_species ~ Land.Use.Description_grouped , data=Plots)
anova(fitSpecies)          #SOMETHING SIGNIFICANT!!!!!!!!!!

# Mortality
fit.mort  = aov(removed_trees.2014 ~ Land.Use.Description_grouped , data=Plots)
anova(fit.mort)          #NOTHING SIGNIFICANT

#======================================================================================
#BA and Biomass correlation with Groundcover
#Dataframe preparation
corr_table = data.frame(matrix(NA, nrow = 8, ncol = 6))
rownames(corr_table) = c('Building',
                         'Cement',
                         'Bare.soil',
                         'Tar',
                         'Herbs',
                         'Grass',
                         'Impermeable',
                         'Permeable' )
colnames(corr_table) = c('BA2002', 
                         'BA2014', 
                         'BA.02.14', 
                         'Biomass.2002',
                         'Biomass.2014', 
                         'Biomass.02.14')

for (response in colnames(corr_table)){
  #png(paste(factor, '.png'))
  for (factor in rownames(corr_table)){
    corr_table[factor,response] = round(cor(Plots[factor], Plots[response])[1], 
                                         digits = 3)
    }
} 

#Correlation coefficients for land cover percentage and BA, Biomass reponses
corr_table
```