---
title:"Santiago_Prepare Dataframes"
author:"Sebastian Palmas"
date:`r format(Sys.time(), '%d %B, %Y')`"
---

This script prepares the Santiago data by importing original database, 

##Get Data
```{r Reading Data, message=FALSE, warning=FALSE}
#Load Packages
require(dplyr)
require(gdata)

# Getting working directory
if (Sys.info()['sysname'] == 'Windows'){
  setwd("C:/Users/Sebastian/Google Drive/Urban")
} else if (Sys.info()['sysname'] == 'Linux'){
  setwd("~/Google Drive/Urban")
}

#  Importing biomass calculating function from helper file
source("Analysis/helpers.R")   

#Reading oriiginal data
arboles <- read.xls(xls="Data/arboles_2002-2014.xls", sheet = 'Parcelas con arboles' )

#Species parameters dataframe
species <- read.csv("Data/species.csv")
rownames(species) <- species$Specie

#Coordinates, Land use ID on MAP, Date sampled, crew, address, groundcovers from PlotGrndCvr.xls
Plots <- read.csv("Data/Plots_original.csv")
rownames(Plots) <- Plots$Plot.ID
#Land use ID used on table above. Description and Abbrev. Sumn of area sampled by Land use
LandUseID <- read.csv("Data/PlotLandUSe_Code.csv")
rownames(LandUseID ) <- LandUseID$Land.Use.ID

#Plot data for 2002
Plots.02 <- read.xls(xls="Data/Stgoplts_2002.xls", sheet = 'Sheet1' )

#Description of shrubs in each plot. 
#Shrubs <- read.xls(xls="Datby_tailnum <- group_by(flights, tailnum)by_tailnum <- group_by(flights, tailnum)a/Shrubs.xls", sheet = 'Shrubs')

#Percent of tree, shrubs and plantables in each plot in 2014
TreeShrbCvr <- read.xls(xls="Data/TreeShrbCvr.xls", sheet = 'Subplots')

#detailed measurements of each tree in 2014 plots. Height to Crown. Tree height, % shrub below
#TreeBiom <- read.xls(xls="Data/TreeBiometrics.xls", sheet = 'Trees')

#percent of plot used in the different classes. ORMITP?
#FieldLandUses <- read.xls(xls="Data/FieldLandUses.xls", sheet = 'FieldLandUses')
```

##Biomass estimation and definition of allometric equations
These are the equation forms that are present in our species. Each species is assigned an equation form and parameters. This data is in species.csv file. This is a way reduce size of database, have only one table with all species information and reduce errors in calculating biomass (avoid calculating with different formulas the same species).  Also makes easy to change equations if errors are found.

1: $\exp(a + b \log(\mathrm{DBH}))$

2: $a + b\mathrm{DBH} + c\mathrm{DBH}^d + e\mathrm{H}^f + g\mathrm{DBH}^2\mathrm{H}$

3: $a \pi(\mathrm{Crown} / 2)^2 \times b \mathrm{H}^c$

4: $a\mathrm{H}^b(\mathrm{DBH}/c)^d$

5: $a\mathrm{DBH}^b+c\mathrm{DBH}^d$

6: $\exp(a+b\mathrm{DBH})$

7: $a+b\log(\mathrm{DBH})$

8: $a\mathrm{H}^b+c\mathrm{H}^d$

0: No Formula

####Data preparation
```{r Data preparation, message=FALSE, warning=FALSE}
# Getting working directory
if (Sys.info()['sysname'] == 'Windows'){
  setwd("C:/Users/Sebastian/Google Drive/Urban")
} else if (Sys.info()['sysname'] == 'Linux'){
  setwd("~/Google Drive/Urban")
}


#Biomass estimation from formulas using helper function
arboles$Biomass.2002 = biomass(spp = arboles$Specie,
                               dbh =arboles$DBH2002,
                               h = arboles$H2002,
                               crown = arboles$CROWN2002)
arboles$Biomass.2014 = biomass(spp = arboles$Specie,
                               dbh =arboles$DBH2014, 
                               h = arboles$H2014,
                               crown = arboles$CROWN2014)


#Change to NA the values of DBH2014, H2014 and CROWN2014 from the removed trees based on Removed.2014 == 1
arboles$DBH2014[arboles$Removed.2014==1]  = NA
arboles$H2014[arboles$Removed.2014==1]  = NA
arboles$CROWN2014[arboles$Removed.2014==1]  = NA
arboles$Biomass.2014[arboles$Removed.2014==1]  = 0
arboles$Differences.in.growth = NULL
arboles$Biomass14.02 = NULL  #Eliminates previously biomass estimates
arboles$DBH2002b = NULL #Eliminates weird second dbh
arboles$DBH2014b = NULL #Eliminates weird second dbh

#Change to NA the values of DBH2002, H2002 and CROWN2002 from the ingrowth trees based on In.growth.2014 == 1
arboles$DBH2002[arboles$In.growth.2014==1]  = NA
arboles$H2002[arboles$In.growth.2014==1]  = NA
arboles$CROWN2002[arboles$In.growth.2014==1]  = NA
arboles$Biomass.2002[arboles$In.growth.2014==1]  = 0 #Serves to properly estimat e change later.

#The change in DBH column only has the trees with measurments in both times. #Only for growth regression
#Considering a linear dbh growth
arboles$DBH.02.12 = (arboles$DBH2014 - arboles$DBH2002) / arboles$Days.between.sampling.dates * (365 * 10)
arboles$DBH.02.12[arboles$In.growth.2014==1]  = NA 

#Estimate basal area in square meters
arboles$BA2002 = (arboles$DBH2002/100/2)^2 * pi
arboles$BA2014 = (arboles$DBH2014/100/2)^2 * pi
#The change in basal area adds the basal area of the recruited trees.
#considering a linear basal area growht (not estimated with adjusted dbh)
arboles$BA2002[arboles$In.growth.2014==1]  = 0
arboles$BA2014[arboles$Removed.2014==1]  = 0
arboles$BA.02.12 = arboles$BA2014 - arboles$BA2002 / arboles$Days.between.sampling.dates * (365 * 10)
arboles$BA.02.14 = arboles$BA2014 - arboles$BA2002
arboles$BA2002[arboles$In.growth.2014 == 1] = NA
arboles$BA2014[arboles$Removed.2014==1]  = NA

#The change in biomass adds the basal area of the recruited trees.
arboles$Biomass.02.14 = arboles$Biomass.2014 - arboles$Biomass.2002
arboles$Biomass.2002[arboles$In.growth.2014==1] = NA    #trees recruited in 2014 do not add to the biomass in 2002. 
arboles$Biomass.2014[arboles$Removed.2014==1]  = NA   #trees dead by 2014 do not add to the biomass in 2014. 

#Adding Plot land use based on PLot ID to arboles list
arboles$Land.Use.ID = Plots[as.character(arboles$Plot.ID),'Land.Use.ID.on.Map']
arboles$Land.Use.Description = LandUseID[as.character(arboles$Land.Use.ID),'LandUseAbbreviation']
arboles$Land.Use.Description_grouped = LandUseID[as.character(arboles$Land.Use.ID),'LandUseAbbreviation_grouped']
arboles$GrndCvrClass = Plots[as.character(arboles$Land.Use.ID),'Perm_degree']

#Add species code and classification of trees from excel
arboles$Class = species[as.character(arboles$Specie),'Class']
#arboles$Specie_code = species[as.character(arboles$Specie),'Specie_code']

#EXporting database of only trees found in both years to estimate growth
write.csv(arboles[!is.na(arboles$DBH.02.12),] , file = 'Data/arboles_2002-2014_growth.csv', row.names = FALSE)
#EXporting database of all trees, now with data prepared (Same trees as arboles_2002-2014.xls)
write.csv(arboles , file = 'Data/arboles_2002-2014_prepared.csv', row.names = FALSE)
```

#### By Plot statistics table export
```{r Plot Statistics, message=FALSE, warning=FALSE}
require(dplyr)

# Getting working directory
if (Sys.info()['sysname'] == 'Windows'){
  setwd("C:/Users/Sebastian/Google Drive/Urban")
} else if (Sys.info()['sysname'] == 'Linux'){
  setwd("~/Google Drive/Urban")
}

by_plot <- dplyr::group_by(arboles, Plot.ID)
by_plot_results = summarise (by_plot,
                             GrndCvrClass  = unique(GrndCvrClass),   #the ground cover class of the parcela
                             num_species = length(unique(Specie)),   #the number of unique species in the parcela
                             num_trees_2002 = length(Specie) - sum(In.growth.2014, na.rm = TRUE),   #the number of trees found in 2002
                             num_trees_2014 = length(Specie) - sum(Removed.2014, na.rm = TRUE),    #the number of trees found in 2014
                             removed_trees.2014 = sum(Removed.2014, na.rm = TRUE),     #the number of deaths from 2002 to 2014
                             ingrowth_trees.2014 = sum(In.growth.2014, na.rm = TRUE),    #the number of new ntrees appeared in 2014
                             #Biomass is also changed to per hectare bases by multiplying by 25 the 400 square meter plots
                             BA2002 = sum(BA2002, na.rm = TRUE)*25,        #Total basal area in 2002
                             BA2014 = sum(BA2014, na.rm = TRUE)*25,        #Total Basal area in 2014
                             BA.02.14 = sum(BA.02.14, na.rm = TRUE)*25,    #change of basal area from 2002-2014
                             #biomass is *25 because these are 400 m2 plots and we need to convert to 10000 for per hectare values
                             #/10000 because it is easier to see in megagrams of carbon
                             Biomass.2002 = sum(Biomass.2002, na.rm = TRUE)*25/1000,    #Total biomass in 2002
                             Biomass.2014 = sum(Biomass.2014, na.rm = TRUE)*25/1000,    #Total Biomass area in 2014
                             Biomass.02.14 = sum(Biomass.02.14, na.rm = TRUE)*25/1000    #change of biomass from 2002-2014
                             )


#plots ya tiene las columnas aqui? calculadas. 
#Hay que eliminarlas antes si se quiere exportar de nuevo
#Adding add these results to the original Plots dataframe and export it for later analysis
Plots = merge(Plots, by_plot_results, by = 'Plot.ID',
              all = TRUE)     #Even if there is no matching, export the line
#Export Plots dataframe to speed later analysis

#--------------Adding Tree % Cover to database
#2002 Data
Plots.02.order <- match(Plots$Plot.ID, Plots.02$PLOT.I.D.) #Getting positions of the PlotsID in the Plots.02 table
Plots['TreeCover02'] <- Plots.02$X.TREE2[Plots.02.order]

TreeShrbCvr.order <- match(Plots$Plot.ID, TreeShrbCvr$PlotID)
Plots['TreeCover14'] <- TreeShrbCvr$PercentTreeCover[TreeShrbCvr.order]

#Creating a column of the difference 2014-2002
Plots['TreeCover.02.14'] <- Plots['TreeCover14']  - Plots['TreeCover02']

#------Adding Land.Use.Description and Land.Use.Description and Land.Use.Description_grouped
Plots$Land.Use.Description = LandUseID[as.character(Plots$Land.Use.ID),'LandUseAbbreviation']
Plots$Land.Use.Description_grouped = LandUseID[as.character(Plots$Land.Use.ID),'LandUseAbbreviation_grouped']
#arboles$GrndCvrClass = Plots[as.character(arboles$Land.Use.ID),'Perm_degree']


#Convert NA to zeroes for later analysos
Plots$Land.Use.Description[is.na(Plots$Land.Use.Description)] <- 0
Plots$Land.Use.Description_grouped[is.na(Plots$Land.Use.Description_grouped)] <- 0
Plots$GrndCvrClass[is.na(Plots$GrndCvrClass)] <- 0
Plots$num_species[is.na(Plots$num_species)] <- 0
Plots$num_trees_2002[is.na(Plots$num_trees_2002)] <- 0
Plots$num_trees_2014[is.na(Plots$num_trees_2014)] <- 0
Plots$removed_trees.2014[is.na(Plots$removed_trees.2014)] <- 0
Plots$ingrowth_trees.2014[is.na(Plots$ingrowth_trees.2014)] <- 0
Plots$BA2002[is.na(Plots$BA2002)] <- 0
Plots$BA2014[is.na(Plots$BA2014)] <- 0
Plots$Biomass.2002[is.na(Plots$Biomass.2002)] <- 0
Plots$BA.02.14[is.na(Plots$BA.02.14)] <- 0
Plots$Biomass.02.14[is.na(Plots$Biomass.02.14)] <- 0
Plots$Biomass.2014[is.na(Plots$Biomass.2014)] <- 0

#Having a column og log biomass
Plots['logBiomass.2014'] <- log(Plots$Biomass.2014)

#Changing the Inf value of the log to  NA
Plots$logBiomass.2014[is.infinite(Plots$logBiomass.2014)] <- NA   #Changing Infinity values of log Biomass to NA


#Export plots analysis to csv file
write.csv(Plots, file = 'Data/Plots.csv', row.names = FALSE)
```

