<<<<<<< HEAD
---
title: "Analysis of Santiago Urban Tree by Land Use"
author: "spp"
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

Script to fit Biomass model from Urban trees in Santiago.

## reding data
```{r Reading Results Data, eval=TRUE, message=FALSE, warnings = FALSE}
require(ggplot2)
require(gplots)
require(multcomp)

#Set working directory
if (Sys.info()['sysname'] == 'Windows'){
  setwd("C:/Users/Sebastian/Google Drive/Urban")
} else if (Sys.info()['sysname'] == 'Linux'){
  setwd("~/Google Drive/Urban")
}

#Reading Data
Plots <- read.csv("Data/Plots.csv")
```


## Correlation between BA and Biomass with estrato, comuna, lat/long
```{r Comuna analysis, eval=TRUE, message=FALSE, warnings = FALSE, echo = FALSE}

```


## By LandUse statistical analysis and plots
```{r by Land Use analysis, eval=TRUE, message=FALSE, warnings = FALSE}
#print number of plots per land use (After combining the many land usesws)
table(Plots$Land.Use.Description_grouped)

#======================================================================================
# Anova of BA change, biomass change by land use (R, C/I, G, A, T)
#with log of BA and Bio


# 2012-2014 Change in Basal Area
fitBA_change <- aov(BA.02.14 ~ Land.Use.Description_grouped , data=Plots)
anova(fitBA_change)          #NOTHING SIGNIFICANT

# 2012-2014 Change in log(Basal Area)
fit_log_BA_change <- aov(log(BA.02.14 + 1) ~ Land.Use.Description_grouped , data=Plots)
anova(fit_log_BA_change)          #NOTHING SIGNIFICANT

# 2012-2014 Change in Biomass
fitBiomass_change <- aov(Biomass.02.14 ~ Land.Use.Description_grouped , data=Plots)
anova(fitBiomass_change)          #NOTHING SIGNIFICANT

# 2012-2014 Change in log(Biomass)
fit_log_Biomass_change <- aov(log(Biomass.02.14 + 1) ~ Land.Use.Description_grouped , data=Plots)
anova(fit_log_Biomass_change)          #NOTHING SIGNIFICANT

#======================================================================================
# Anova of BA, biomass, number of species, mortality
# by grouped land use Classification (R, C/I, G, A, T)

# 2002 Basal Area
fitBA.2002  = aov(BA2002 ~ Land.Use.Description_grouped , data=Plots)
anova(fitBA.2002)          #NOTHING SIGNIFICANT

# 2014 Basal Area
fitBA.2014  = aov(BA2014 ~ Land.Use.Description_grouped , data=Plots)
anova(fitBA.2014)          #NOTHING SIGNIFICANT

# 2002 Biomass
fitBiomass.2002  = aov(Biomass.2002 ~ Land.Use.Description_grouped , data=Plots)
anova(fitBiomass.2002)          #NOTHING SIGNIFICANT

# 2014 Basal Area
fitBiomass.2014  = aov(Biomass.2014 ~ Land.Use.Description_grouped , data=Plots)
anova(fitBiomass.2014)          #NOTHING SIGNIFICANT

# Total number of species
fitSpecies  <- aov(num_species ~ Land.Use.Description_grouped , data=Plots)
anova(fitSpecies)          #SOMETHING SIGNIFICANT!!!!!!!!!!

# Mortality
fit.mort  = aov(removed_trees.2014 ~ Land.Use.Description_grouped , data=Plots)
anova(fit.mort)          #NOTHING SIGNIFICANT

# Resprouting
fit.resp  = aov(ingrowth_trees.2014 ~ Land.Use.Description_grouped , data=Plots)
anova(fit.resp)          #NOTHING SIGNIFICANT
```



## Correlation between BA and Biomass with goundcover percentages
```{r correlation with Groundcover, eval=TRUE, message=FALSE, warnings = FALSE, echo = FALSE}
#BA and Biomass 
#Dataframe preparation
corr_table = data.frame(matrix(NA, nrow = 8, ncol = 6))
rownames(corr_table) = c('Building',
                         'Cement',
                         'Bare.soil',
                         'Tar',
                         'Herbs',
                         'Grass',
                         'Impermeable',
                         'Permeable' )
colnames(corr_table) = c('BA2002', 
                         'BA2014', 
                         'BA.02.14', 
                         'Biomass.2002',
                         'Biomass.2014', 
                         'Biomass.02.14')

for (response in colnames(corr_table)){
  #png(paste(factor, '.png'))
  for (factor in rownames(corr_table)){
    corr_table[factor,response] = round(cor(Plots[factor], Plots[response], use = 'complete.obs')[1], 
                                         digits = 3)
    }
} 
#Correlation coefficients for land cover percentage and BA, Biomass reponses
corr_table
```

=======
---
title: "Analysis of Santiago Urban Tree by Land Use"
author: "spp"
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

Script to fit Biomass model from Urban trees in Santiago.



## By Plots statistical analysis and plots
```{r Reading Results Data, eval=TRUE, message=FALSE, warnings = FALSE}
require(ggplot2)
require(gplots)
require(multcomp)

#Set working directory
if (Sys.info()['sysname'] == 'Windows'){
  setwd("C:/Users/Sebastian/Google Drive/Urban")
} else if (Sys.info()['sysname'] == 'Linux'){
  setwd("~/Google Drive/Urban")
}

#Reading Data
Plots <- read.csv("Data/Plots.csv")

#print number of plots per land use (After combining the many land usesws)
table(Plots$Land.Use.Description_grouped)

#======================================================================================
# Anova of BA change, biomass change by land use (R, C/I, G, A, T)

# 2012-2014 Change in Basal Area
fitBA_change <- aov(BA.02.14 ~ Land.Use.Description_grouped , data=Plots)
anova(fitBA_change)          #NOTHING SIGNIFICANT

# 2012-2014 Change in log(Basal Area)
fit_log_BA_change <- aov(log(BA.02.14 + 1) ~ Land.Use.Description_grouped , data=Plots)
anova(fit_log_BA_change)          #NOTHING SIGNIFICANT

# 2012-2014 Change in Biomass
fitBiomass_change <- aov(Biomass.02.14 ~ Land.Use.Description_grouped , data=Plots)
anova(fitBiomass_change)          #NOTHING SIGNIFICANT

# 2012-2014 Change in log(Biomass)
fit_log_Biomass_change <- aov(log(Biomass.02.14 + 1) ~ Land.Use.Description_grouped , data=Plots)
anova(fit_log_Biomass_change)          #NOTHING SIGNIFICANT

#======================================================================================
# Anova of BA, biomass, number of species, mortality
# by grouped land use Classification (R, C/I, G, A, T)

# 2002 Basal Area
fitBA.2002  = aov(BA2002 ~ Land.Use.Description_grouped , data=Plots)
anova(fitBA.2002)          #NOTHING SIGNIFICANT

# 2014 Basal Area
fitBA.2014  = aov(BA2014 ~ Land.Use.Description_grouped , data=Plots)
anova(fitBA.2014)          #NOTHING SIGNIFICANT

# 2002 Biomass
fitBiomass.2002  = aov(Biomass.2002 ~ Land.Use.Description_grouped , data=Plots)
anova(fitBiomass.2002)          #NOTHING SIGNIFICANT

# 2014 Basal Area
fitBiomass.2014  = aov(Biomass.2014 ~ Land.Use.Description_grouped , data=Plots)
anova(fitBiomass.2014)          #NOTHING SIGNIFICANT

# Total number of species
fitSpecies  <- aov(num_species ~ Land.Use.Description_grouped , data=Plots)
anova(fitSpecies)          #SOMETHING SIGNIFICANT!!!!!!!!!!

# Mortality
fit.mort  = aov(removed_trees.2014 ~ Land.Use.Description_grouped , data=Plots)
anova(fit.mort)          #NOTHING SIGNIFICANT

# Resprouting
fit.resp  = aov(ingrowth_trees.2014 ~ Land.Use.Description_grouped , data=Plots)
anova(fit.resp)          #NOTHING SIGNIFICANT

#======================================================================================
#BA and Biomass correlation with Groundcover
#Dataframe preparation
corr_table = data.frame(matrix(NA, nrow = 8, ncol = 6))
rownames(corr_table) = c('Building',
                         'Cement',
                         'Bare.soil',
                         'Tar',
                         'Herbs',
                         'Grass',
                         'Impermeable',
                         'Permeable' )
colnames(corr_table) = c('BA2002', 
                         'BA2014', 
                         'BA.02.14', 
                         'Biomass.2002',
                         'Biomass.2014', 
                         'Biomass.02.14')

for (response in colnames(corr_table)){
  #png(paste(factor, '.png'))
  for (factor in rownames(corr_table)){
    corr_table[factor,response] = round(cor(Plots[factor], Plots[response], use = 'complete.obs')[1], 
                                         digits = 3)
    }
} 

#Correlation coefficients for land cover percentage and BA, Biomass reponses
corr_table
```

>>>>>>> 5c4efa2c7ce696bd54de300cccccda4d481dbe6c
